# Creating tables in huxtable

[Huxtable](https://hughjonesd.github.io/huxtable/) is an R library to create tables that can be output to LaTeX, HTML, and Word documents.



## Requirements

- To use huxtable requires [pandoc](https://pandoc.org/) version [>2.0](https://github.com/jgm/pandoc/releases/tag/2.8.0.1). 
If you can't update the system pandoc you can set the PATH variable to point to a local version when running R. 


- To output to word document huxtable also requires [flextable](https://davidgohel.github.io/flextable/index.html)



## Example use

In this example we are going to use the following toy data frame.

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
example.df <- data.frame(Col1=c('Foo', 'Bar', 'Fizz', 'Buzz'),
                         Col2=c(1, 2, 3, 4),
                         Col3=c(1.12345, 1.54321, 4.12345, 4.54321))

```

The first thing we need to do is turn the dataframe `example.df` into a huxtable.


### Default output

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(huxtable)

example.huxtable <- as_hux(example.df)
```

The huxtable `example.huxtable` by default looks like this.
```{r echo = FALSE, results='asis', warning=FALSE, cache=FALSE}
example.huxtable
```

### Adding column names to the table

We can add formatting to this table. Here we use the pipe interface to huxtable. 
Using the pipe interface requires the `magrittr` library. 

The first thing we want to do is add the column names to the outputted table.
To do this we pipe `example.huxtable` through `add_colnames`. 

The huxtable `example.huxtable` by default looks like this.
```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

example.huxtable %>% 
add_colnames
```

