# Creating tables in huxtable

[Huxtable](https://hughjonesd.github.io/huxtable/) is an R library to create tables that can be output to LaTeX, HTML, and Word documents.



## Requirements

- To use huxtable requires [pandoc](https://pandoc.org/) version [>2.0](https://github.com/jgm/pandoc/releases/tag/2.8.0.1). 
If you can't update the system pandoc you can set the PATH variable to point to a local version when running R. 


- To output to word document huxtable also requires [flextable](https://davidgohel.github.io/flextable/index.html)

## Caveats

The huxtable package doesn't support captions for word documents.


## Example use

In this example we are going to use the following toy data frame.

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
example.df <- data.frame(Col1=c('Foo', 'Bar', 'Fizz', 'Buzz'),
                         Col2=c(1, 2, 3, 4),
                         Col3=c(1.12345, 1.54321, 4.12345, 4.54321))

```

The first thing we need to do is turn the dataframe `example.df` into a huxtable.


### Default output

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(huxtable)

example.huxtable <- as_hux(example.df)
```

The huxtable `example.huxtable` by default looks like this.
```{r echo = FALSE, results='asis', warning=FALSE, cache=FALSE}
example.huxtable
```

### Adding column names to the table

We can add formatting to this table. Here we use the pipe interface to huxtable. 
Using the pipe interface requires the `magrittr` library. 

The first thing we want to do is add the column names to the outputted table.
To do this we pipe `example.huxtable` through `add_colnames`. 

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

example.huxtable %>% 
add_colnames
```

### Bolding cells

Next we add bolding to the 1st row by piping to `set_bold`.
The first argument of the pipe interface is the row number to apply bold to.
The second argument is the column number where `everywhere` means every column. 
The third argument is a boolean value for whether the cell is bold (`TRUE`) or not (`FALSE`).


```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

example.huxtable %>% 
add_colnames %>%
set_bold(1, everywhere, TRUE)
```

### Cell borders

To add a bottom border to the first column we use `set_bottom_border` providing the row, columns, and line width of the border.
Similarly we use `set_top_border`.

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

example.huxtable %>% 
add_colnames %>%
set_bold(1, everywhere, TRUE) %>%
set_bottom_border(1, everywhere, 2) %>%
set_top_border(1, everywhere, 2)
```


### Adding rows and merging cells

We can add rows using `add_rows` and then we can merge cells in a row using `merge_cells`.

```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

hdr1 <- hux("Sub-header for FooBar", "", "")
hdr2 <- hux("Sub-header for FizzBuzz", "", "")

example.huxtable %>% 
add_colnames %>%
set_bold(1, everywhere, TRUE) %>%
set_bottom_border(1, everywhere, 2) %>%
set_top_border(1, everywhere, 2) %>%
add_rows(hdr1, 1) %>%  # the row is added **after** the specified row (new row is row 2)
merge_cells(2, c(1, 3)) %>%
add_rows(hdr2, 4) %>%  # at this point "Bar" is in row 4
merge_cells(5, c(1, 3))

```

### Adding footnotes

To add footnotes to a huxtable can be done by piping to `add_footnote`.


```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

hdr1 <- hux("Sub-header for FooBar", "", "")
hdr2 <- hux("Sub-header for FizzBuzz", "", "")

example.huxtable %>% 
add_colnames %>%
set_bold(1, everywhere, TRUE) %>%
set_bottom_border(1, everywhere, 2) %>%
set_top_border(1, everywhere, 2) %>%
add_rows(hdr1, 1) %>%  # the row is added **after** the specified row (new row is row 2)
merge_cells(2, c(1, 3)) %>%
add_rows(hdr2, 4) %>%  # at this point "Bar" is in row 4
merge_cells(5, c(1, 3)) %>%
add_footnote("This is the footnote text for the table")

```


### Adding a caption

A caption can be added using `set_caption` and the position of the caption can be set using `set_caption_pos`.
However this will only work when exporting to LaTeX, HTML, and PDF.


```{r echo = TRUE, results='asis', warning=FALSE, cache=FALSE}
library(magrittr)

hdr1 <- hux("Sub-header for FooBar", "", "")
hdr2 <- hux("Sub-header for FizzBuzz", "", "")

example.huxtable %>% 
add_colnames %>%
set_bold(1, everywhere, TRUE) %>%
set_bottom_border(1, everywhere, 2) %>%
set_top_border(1, everywhere, 2) %>%
add_rows(hdr1, 1) %>%  # the row is added **after** the specified row (new row is row 2)
merge_cells(2, c(1, 3)) %>%
add_rows(hdr2, 4) %>%  # at this point "Bar" is in row 4
merge_cells(5, c(1, 3)) %>%
add_footnote("This is the footnote text for the table") %>%
set_caption("Caption text for table") %>%
set_caption_pos("bottom")

```





